---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# dmhct

<!-- badges: start -->
[![R-CMD-check](https://github.com/jesse-smith/dmhct/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/jesse-smith/dmhct/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`{dmhct}` extracts, loads, and transforms data for the MLinHCT project. The main
functionality is wrapped into a single function, `dm_elt()`, that produces a
`dm` object (from the [`{dm}` package](https://cynkra.github.io/dm/)) containing
cleaned data tables from an SQL Server database.

## Installation and Setup

You can install `{dmhct}` from Github with:

``` r
# install.packages("remotes")
remotes::install_github("jesse-smith/dmhct")
```

To use the package, you will need access to the SQL Server instance containing
the data. This is provided by St. Jude IT; contact
[Jesse Smith](mailto:jesse.smith@stjude.org) to learn how to gain access.


## Introduction

The main functionality in the package is in `dm_elt()`; this function combines
two sub-functions, `dm_extract()` and `dm_transform()`, into a small pipeline.

```{r example}
library(dmhct)
dm <- dm_elt()
dm2 <- dm_extract() %>% dm_transform()
identical(dm, dm2)
```

```{r delete-dm2, include=FALSE}
rm(dm2)
```

The `dm_extract()` function is itself a wrapper for a pipeline used to extract
the necessary data from the SQL Server and load it onto your local machine. The
`dm_transform()` function provides a second pipeline that cleans each table in
the database and outputs tables ready for analysis
(as [`data.table`s](https://rdatatable.gitlab.io/data.table/)).

Both functions cache their results by default and compare a checksum of their
inputs to a cached checksum to determine whether re-calculation is necessary.
This allows the functions to simply read the cached outputs, rather than repeat
costly computations, if the input is unchanged. Each function also allows the
user to force re-computation and cache overwriting via a `reset` argument. This
caching behavior (and the control arguments) are made available in the `dm_elt()`
wrapper as well.

## Tables

The data used in the project consists of `r length(dm)` tables:

```{r dm-summary}
summary(dm)[, -3L]
```

The entry point to this data is the `master` table, which contains one row per
patient and includes baseline variables such as transplant date,
patient demographics, and donor characteristics. It also contains unique IDs for
patients (`entity_id`) and donors (`donor_id`).

```{r master-summary}
skimr::skim(dm$master, setdiff(colnames(dm$master), c("entity_id", "donor_id")))
```

All tables are keyed by one or more columns that are (jointly) unique to each
row. Each table contains `entity_id` as a primary key; most tables except `master`
then define additional keys specific to the data contained in that table (for
example, many tables contain repeated observations for a patient and thus add a
`date` key).

```{dm-pks}
dm::dm_get_all_pks(dm) %>% dplyr::mutate(pk_col = format(.data$pk_col))
```

Here's a list of data (non-key) columns in each table (excluding `master`):

```{r dm-data}
tibble::tibble(
  table = names(dm::dm_select_tbl(dm, -"master")),
  cols = sapply(dm::dm_select_tbl(dm, -"master"), function(x) {
    paste0(setdiff(colnames(x), data.table::key(x)), collapse = ", ")
  })
)
```

## Development

You can find code generating results for each table in a corresponding file
named `"dm_[table-name]"`. Changes should not be committed directly to the `main`
branch; going forward, all changes should be made in a feature/bug fix/etc. branch
and reviewed by another team member before merging. The Github repository is set
up to dis-allow pushing direct commits to `main`.

## Using with `{predHCT}`

`{dmhct}` was created to de-couple data cleaning code from modeling code for the
MLinHCT project. However, it is a required dependency for the main modeling package,
`{predHCT}`. While it is possible to install your local development version for
use with `{predHCT}`, this is not recommended. Instead, please install from the
`main` Github branch using the installation instructions at the beginning of
this document.
